---
title: "Homework 2"
output: github_document
---

This is my solution for Homework 2. 

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
```


## Problem 1

Read the Mr. Trash Wheel dataset. 

```{r trashwheel_data}
trashwheel_df = 
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel", 
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2017 and 2018. 

```{r precipitation_data}
precip_2018 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation", 
  skip = 1, 
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation", 
  skip = 1, 
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Combine annual precipitation. 

```{r annual_precipitation}
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows by `r ncol(trashwheel_df)` columns in our final dataset. Additional data sheets include monthly precipitation data. The total precipitation in 2019 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. The median number of sports balls found in the a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`. 

## Problem 2

Read and clean the NYC Transit dataset. 

```{r read_transit_data}
transit_df = 
  read.csv("./NYC_Transit_Subway_Entrance_And_Exit_Data (1).csv") %>%
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) 
  mutate(transit_df, entry = recode(entry, YES = "TRUE", NO = "FALSE"))
```



```{r describing_transit_data}
select(distinct(transit_df, line, station_name)) %>% 
  count()

filter(transit_df, ada == "TRUE") %>%
    distinct(line, station_name) %>%
    count()

filter(transit_df, vending == "NO" & entry == "YES") %>% 
  count() / 
filter(transit_df, vending == "NO") %>% 
      count() 
```

This dataset contains information about the NYC transit system. It contains the variables: `r variable.names(transit_df)`. The data cleaning so far has consisted of cleaning the variable names, selecting appropriate columns in the dataset, and converting the entry variable from a character to logical variable. The resulting dataset is `r nrow(transit_df)` rows by `r ncol(transit_df)` columns. 

There are `r select(distinct(transit_df, line, station_name)) %>% count()` distinct stations. `r filter(transit_df, ada == "TRUE") %>% distinct(line, station_name) %>% count()` of the stations are ADA compliant and `r filter(transit_df, vending == "NO" & entry == "YES") %>% count()` out of `r filter(transit_df, vending == "NO") %>% count()` station entrance/exits without vending allow entrance. 

```{r transit_tidy}
transit_tidy =
transit_df %>%
mutate_at(vars(route1:route11), as.character) %>%
pivot_longer(
route1:route11,
names_to = "route_name",
names_prefix = "route",
values_to = "route_number") %>%
drop_na(route_number)

filter(transit_tidy, route_number == "A") %>% 
  distinct(route_number, station_name) %>% 
  count()

filter(transit_tidy, route_number == "A" & ada == TRUE) %>% 
    distinct(route_number, station_name) %>%
    count()

```

There are `r filter(transit_tidy, route_number == "A") %>% distinct(route_number, station_name) %>% count()` distinct stations that serve the A train. Of these stations, `r filter(transit_tidy, route_number == "A" & ada == TRUE) %>% distinct(route_number, station_name) %>% count()` are ADA compliant. 

## Problem 3 

Read and clean FiveThirtyEight pols-month dataset. 

```{r read_pols_month}
pols_df = 
  read.csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, c("year", "month", "day"), sep="-") %>% 
  mutate(month = month.abb[as.factor(month)]) %>% 
  mutate(president = case_when(prez_gop == 1 ~ "gop", prez_dem == 1 ~ "dem")) %>% 
  select(-day, -prez_dem, -prez_gop)
```

Read and clean snp.csv dataset. 

```{r snp_data}
snp_df = 
read.csv("./fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, c("year", "month", "day"), sep="/") %>% 
  mutate(month = month.abb[as.factor(month)]) %>% 
  select(-day)
```



