---
title: "Homework 2"
output: github_document
---

This is my solution for Homework 2. 

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```


## Problem 1

Read the Mr. Trash Wheel dataset. 

```{r}
trashwheel_df = 
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel", 
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2017 and 2018. 

```{r}
precip_2018 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation", 
  skip = 1, 
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation", 
  skip = 1, 
 ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Combine annual precipitation. 

```{r}
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trash Wheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trash wheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, including some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows by `r ncol(trashwheel_df)` columns in our final dataset. Additional data sheets include monthly precipitation data. The total precipitation in 2019 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches. THe median number of sports balls found in the a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`. 

